Updates a Procrustes and a PCA

### 2025-10-24 Analysis/17-update_procrustes_and_PCA.Rmd
Update 17-update_procrustes_and_PCA.Rmd

## This value needs to be multiplied by 1.048-ish for some reason
prev_mean_shape <- whole_gpa$mshape # * 1.048

for (i in 1:n) {
    newdatarot[,,i]  <- rotonto(x = prev_mean_shape, y = new_landmarks[,,i], scale = atts$scale, reflection = atts$reflect, centerweight = atts$centerweight, weights = atts$weights)$yrot
}

@@

## Getting the new PCAs
new_pca_orpda <- stats::predict(original_pca, newdata = proctan_ordat)
new_pca_norop <- stats::predict(original_pca, newdata = proctan_norop)

@@

plot(whole_gpa$PCscores[,c(1,2)], col = "grey", main = "Sanity check")
points(updated_pca[, c(1,2)], pch = 19, cex = 0.5, col = "grey")
points(whole_gpa$PCscores[dimnames(testing_projection_shapes)[[3]], c(1,2)], pch = 21, col = "darkgreen")
points(new_pca_orpda[, c(1,2)], col = "blue", pch = 19, cex = 0.5)
points(new_pca_norop[, c(1,2)], col = "orange", pch = 19, cex = 0.5)

>>>>

## This value needs to be multiplied by 1.048-ish for some reason or 0.954
prev_mean_shape <- whole_gpa$mshape # / 0.954 # * 1.048

## n Previous specimens = 8748
## n New specimens = 20

## Kind of works.
scalars <- seq(from = 0.9538, to = 0.9552, length.out = 20)

for (i in 1:n) {

    ## New shape should be updated each time?
    prev_mean_shape <- whole_gpa$mshape / scalars[i]
    ## Test solution: update gpa (and mshape) one spec at the time? Like:
    # updated_gpa <- previous_gpa
    # for (i in 1:new_spec) {
    #     updated_gpa <- update.gpa(updated_gpa, new_specimen)
    # }

    newdatarot[,,i]  <- rotonto(x = prev_mean_shape, y = new_landmarks[,,i], scale = atts$scale, reflection = atts$reflect, centerweight = atts$centerweight, weights = atts$weights)$yrot
}

@@

## Getting the new PCAs
new_pca_orpda <- stats::predict(original_pca, newdata = proctan_ordat)
new_pca_norop <- stats::predict(original_pca, newdata = proctan_norop)
        # -> No difference between orp or not

@@

plot(whole_gpa$PCscores[,c(1,2)], col = "grey", main = "Sanity check")
points(updated_pca[, c(1,2)], pch = 19, cex = 0.5, col = "grey")
points(whole_gpa$PCscores[dimnames(testing_projection_shapes)[[3]], c(1,2)], pch = 21, col = "darkgreen")
points(new_pca_orpda[, c(1,2)], col = "blue", pch = 19, cex = 0.5)
# points(new_pca_norop[, c(1,2)], col = "orange", pch = 19, cex = 0.5)

ratios_orp <- new_pca_orpda[, c(1,2)] / whole_gpa$PCscores[dimnames(testing_projection_shapes)[[3]], c(1,2)]
ratios_noorp <- new_pca_norop[, c(1,2)] / whole_gpa$PCscores[dimnames(testing_projection_shapes)[[3]], c(1,2)]

plot(sort(ratios_orp[, 1]), type = "l")
lines(sort(ratios_orp[, 2]), col = "red")

points(new_pca_orpda[, c(1,2)] / ratios, col = "blue", pch = 19, cex = 0.5)
